Q1: Are lake and terrestrial primary productivity coherent?

### 2019-06-14 AnalysisCode/ms1_analyses_randforest.Rmd
random forest partial dependence plots

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)
library(raster)
library(MuMIn)
library(dplyr)
library(nlme)
library(lmerTest)
library(viridis)
library(fields)

@@

```{r modelling, echo=TRUE, cache=FALSE}

library(partykit)

cforest.st<-cforest(accndvicoh.ts1 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi + hu4_zoneid + cv.accndvi, data=modvars.accndvi)
predcoh.st<-predict.cforest(cforest.st, newdata=modvars.accndvi)
hist(predcoh.st)
hist(modvars.accndvi$accndvicoh.ts1)
plot(predcoh.st, modvars.accndvi$accndvicoh.ts1, xlab="predicted", ylab="empirical", main="Coherence, short ts",
     xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1)
cor.test(predcoh.st,modvars.accndvi$accndvicoh.ts1)
varimp.coh.st<-varimp.cforest(cforest.st)
print(varimp.coh.st)

cforest.lt<-cforest(accndvicoh.ts2 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi + hu4_zoneid + cv.accndvi, data=modvars.accndvi)
predcoh.lt<-predict.cforest(cforest.lt, newdata=modvars.accndvi)
hist(predcoh.lt)
hist(modvars.accndvi$accndvicoh.ts2)
plot(predcoh.lt, modvars.accndvi$accndvicoh.ts2, xlab="predicted", ylab="empirical", main="Coherence, long st",
     xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1)
cor.test(predcoh.lt,modvars.accndvi$accndvicoh.ts2)
varimp.coh.lt<-varimp.cforest(cforest.lt)
print(varimp.coh.lt)

cforest.phi.st<-cforest(cos(accndviphi.ts1) ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi + hu4_zoneid + cv.accndvi, 
                        data=modvars.accndvi.phist)
predphi.st<-predict.cforest(cforest.phi.st, newdata=modvars.accndvi.phist)
hist(predphi.st)
hist(cos(modvars.accndvi.phist$accndviphi.ts1))
plot(predphi.st, cos(modvars.accndvi.phist$accndviphi.ts1), xlab="predicted", ylab="empirical", main="cos(phase), short ts",
     xlim=c(-1,1), ylim=c(-1,1))
abline(a=0,b=1)
cor.test(predphi.st,cos(modvars.accndvi.phist$accndviphi.ts1))
varimp.phi.st<-varimp.cforest(cforest.phi.st)
print(varimp.phi.st)

cforest.phi.lt<-cforest(cos(accndviphi.ts2) ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi + hu4_zoneid + cv.accndvi, 
                        data=modvars.accndvi.philt)
predphi.lt<-predict.cforest(cforest.phi.lt, newdata=modvars.accndvi.philt)
hist(predphi.lt)
hist(cos(modvars.accndvi.philt$accndviphi.ts2))
plot(predphi.lt, cos(modvars.accndvi.philt$accndviphi.ts2), xlab="predicted", ylab="empirical", main="cos(phase), short ts",
     xlim=c(-1,1), ylim=c(-1,1))
abline(a=0,b=1)
cor.test(predphi.lt,cos(modvars.accndvi.philt$accndviphi.ts2))
varimp.phi.lt<-varimp.cforest(cforest.phi.lt)
print(varimp.phi.lt)

@@

>>>>

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)
library(raster)
library(MuMIn)
library(dplyr)
library(nlme)
library(lmerTest)
library(viridis)
library(fields)
#library(partykit)
library(pdp)
library(party)

@@

```{r modellingCoherence, echo=TRUE, cache=FALSE}

# cforest.st<-partykit::cforest(accndvicoh.ts1 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + cv.accndvi, 
#                     data=modvars.accndvi, ntree=20000)
cforest.st<-party::cforest(accndvicoh.ts1 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + cv.accndvi, 
                    data=modvars.accndvi, controls=cforest_control(ntree=80000))
predcoh.st<-predict(cforest.st, newdata=modvars.accndvi,type="response")
hist(predcoh.st)
hist(modvars.accndvi$accndvicoh.ts1)
plot(predcoh.st, modvars.accndvi$accndvicoh.ts1, xlab="predicted", ylab="empirical", main="Coherence, short ts",
     xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1)
cor.test(predcoh.st,modvars.accndvi$accndvicoh.ts1)
varimp.coh.st<-varimp(cforest.st)
print(varimp.coh.st[order(varimp.coh.st, decreasing=T)])

partial(cforest.st, pred.var="lake_perim_meters", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="pct.ag", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="lake_area_ha", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="maxdepth", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="chla", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="cv.accndvi", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="tsi.cat", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="hu4_zoneid", train=modvars.accndvi, type="regression", plot=T)

cforest.lt<-party::cforest(accndvicoh.ts2 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + cv.accndvi, 
                    data=modvars.accndvi, controls=cforest_control(ntree=80000))
predcoh.lt<-predict(cforest.lt, newdata=modvars.accndvi)
# hist(predcoh.lt)
# hist(modvars.accndvi$accndvicoh.ts2)
plot(predcoh.lt, modvars.accndvi$accndvicoh.ts2, xlab="predicted", ylab="empirical", main="Coherence, long st",
     xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1)
cor.test(predcoh.lt,modvars.accndvi$accndvicoh.ts2)
varimp.coh.lt<-varimp(cforest.lt)
print(varimp.coh.lt[order(varimp.coh.lt, decreasing=T)])

partial(cforest.lt, pred.var="cv.accndvi", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="hu4_zoneid", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="pct.ag", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="lake_area_ha", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="lake_perim_meters", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="chla", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="tsi.cat", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="maxdepth", train=modvars.accndvi, type="regression", plot=T)

@@

```{r modellingPhase, echo=TRUE, cache=FALSE}

cforest.phi.st<-cforest(cos(accndviphi.ts1) ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + cv.accndvi, 
                        data=modvars.accndvi.phist,controls=cforest_control(ntree=30000))
predphi.st<-predict(cforest.phi.st, newdata=modvars.accndvi.phist)
hist(predphi.st)
hist(cos(modvars.accndvi.phist$accndviphi.ts1))
plot(predphi.st, cos(modvars.accndvi.phist$accndviphi.ts1), xlab="predicted", ylab="empirical", main="cos(phase), short ts",
     xlim=c(-1,1), ylim=c(-1,1))
abline(a=0,b=1)
cor.test(predphi.st,cos(modvars.accndvi.phist$accndviphi.ts1))
varimp.phi.st<-varimp(cforest.phi.st)
print(varimp.phi.st[order(varimp.phi.st,decreasing=TRUE)])

partial(cforest.phi.st, pred.var="cv.accndvi", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="hu4_zoneid", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="pct.ag", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="lake_area_ha", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="lake_perim_meters", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="chla", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="tsi.cat", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="maxdepth", train=modvars.accndvi.phist, type="regression", plot=T)

cforest.phi.lt<-cforest(cos(accndviphi.ts2) ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + cv.accndvi, 
                        data=modvars.accndvi.philt,controls=cforest_control(ntree=30000))
predphi.lt<-predict(cforest.phi.lt, newdata=modvars.accndvi.philt)
hist(predphi.lt)
hist(cos(modvars.accndvi.philt$accndviphi.ts2))
plot(predphi.lt, cos(modvars.accndvi.philt$accndviphi.ts2), xlab="predicted", ylab="empirical", main="cos(phase), short ts",
     xlim=c(-1,1), ylim=c(-1,1))
abline(a=0,b=1)
cor.test(predphi.lt,cos(modvars.accndvi.philt$accndviphi.ts2))
varimp.phi.lt<-varimp(cforest.phi.lt)
print(varimp.phi.lt[order(varimp.phi.lt, decreasing=TRUE)])

partial(cforest.phi.lt, pred.var="hu4_zoneid", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="lake_area_ha", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="lake_perim_meters", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="chla", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="cv.accndvi", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="maxdepth", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="pct.ag", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="tsi.cat", train=modvars.accndvi.philt, type="regression", plot=T)
